%For two sided printing
% \documentclass[a4paper,12pt,twoside,onecolumn,openright,final]{memoir} %for printing two sided

%For one sided printing
\documentclass[a4paper,12pt,oneside,onecolumn,final]{memoir} %for electronic viewing and one-sided printing

%Memoir class geometry, stole it from online manual
\settrimmedsize{297mm}{210mm}{*}
\setlength{\trimtop}{0pt}
\setlength{\trimedge}{\stockwidth}
\addtolength{\trimedge}{-\paperwidth}
\settypeblocksize{634pt}{448.13pt}{*} %original is 634
\setulmargins{4cm}{*}{*} %original is 4 lol
\setlrmargins{*}{*}{1.5}
\setmarginnotes{17pt}{51pt}{\onelineskip}
\setheadfoot{\onelineskip}{2\onelineskip}
\setheaderspaces{*}{2\onelineskip}{*}
\checkandfixthelayout

\usepackage[nodate, 12hr]{datetime} %for time and date
\title{Brief Notes on Differential Geometry}
\author{Lim Kian Hwee}
\date{\small Last updated: \today, \currenttime} %leave blank if dont want date

%Packages for paper formatting, figures, hyperref, etc
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{parskip} %package to tweak paragraph skipping
\usepackage{graphicx}
\usepackage[space]{grffile} %For file names with spaces
\graphicspath{{./Figures/}}
\usepackage{chngcntr}
\counterwithin{figure}{chapter}
\counterwithin{footnote}{chapter}
\usepackage{float}
\usepackage{comment} 
\usepackage{wrapfig}
\setcounter{tocdepth}{1} %stops TOC numbering at sections
\setsecnumdepth{subsection} %numbers subsections
\PassOptionsToPackage{hyphens}{url}
\usepackage[hidelinks]{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={blue!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}
%For quotation marks to behave well
\usepackage{csquotes}
% \MakeOuterQuote{"}

%Blind text
\usepackage[english]{babel}
\usepackage{blindtext}
\blindmathtrue

%Math packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\numberwithin{equation}{chapter} %numbers equations according to chapters
%\allowdisplaybreaks %equations in environment can page break
\usepackage{physics}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\theoremstyle{remark}
\newtheorem{corollary}{Corollary}[theorem]
\theoremstyle{plain}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
% \usepackage{coloremoji}

%Alternate fonts 
\usepackage{newtxtext,newtxmath} %this is good
% \usepackage{fourier} %this is good
% \usepackage{newpxtext,newpxmath} %this is good

%The rest are...meh
% \usepackage{libertine,libertinust1math} %this is quite good
% \usepackage{fouriernc} %this is quite good
% \usepackage{mathptmx} %looks quite bad

%For new look
% \usepackage{newpxtext} \usepackage[euler-digits]{eulervm} %looks disgusting

\begin{document}
  \frontmatter
  \maketitle
  \begin{KeepFromToc}
    \tableofcontents %optional
  \end{KeepFromToc}
  \chapter{Foreword}
    This set of notes is a summary of the NUS Physics module "Mathematical
    Methods in Physics 3" (or MM3 for short) taken in my fourth year of study
    in NUS. Naturally, the references for this set of notes would be:
    \begin{enumerate}
      \item{PC4274 Lecture Notes and tutorials, Kuldip}
      \item{PC4274 Lecture Notes, Edward Teo (to a small extent)}
    \end{enumerate}
    The purpose of this set of notes is to summarise the key ideas of
    Differential Geometry as I understand it, so that in the future I can
    quickly reference this to get a brief idea of the topic. For more
    detailed treatments, there are textbooks like Frankel, Nakahara, or just
    the lecture notes themselves that are listed above. Unless absolutely
    necessary, no examples will be given.

    Also, Einsten summation convention will be used throughout. I.e, for
    tensors, repeated upper and lower indices are summed over. Sometimes we
    will sum over repeated indices for things that are not tensors, in which case we might sometimes sum over two repeated upper indices, or two repeated lower indices (sloppy notation, I know).

    Lastly, most importantly, \emph{Soli Deo Gloria!}.

  \mainmatter
    \include{chapterOneBasicTopology}
    \include{chapterTwoManifoldsAndCoordinates}
    \include{chapterThreeTangentSpaces}
    \include{chapterFourVectorFieldsAndIntegralCurves}
    \include{chapterFiveCotangentSpacesAndOneForms}
    \include{chapterSixTensorsAndTensorFields}
    \include{chapterSevenLieDerivatives}
    % \include{chapterEightnForms}
    % \include{chapterNineExteriorDifferentiation}
    % \include{chapterTenMetricTensor}
    % \include{chapterElevenExteriorCalculus}
    \include{thingsToAdd}
  
\end{document}